---
title: "Day 9 Interactive"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(patchwork)
library(kableExtra)
library(palmerpenguins)
library(reprex)
```

## Warm Up

Starting from penguins, create new subset in which you:

1.  only keep columns species, island, sex, body_mass_g, and flipper_length_mm
2.  filter to only include chinstrap and gentoos
3.  rename species column to penguin_species
4.  create a summary table that contains the mean of body mass, mean of flipper length, and total count of observations each by species and sex

```{r}
penguins %>% 
  select(-year, -bill_depth_mm, -bill_length_mm) %>% 
  filter(species != "Adelie") %>% 
  rename(penguin_species = species) %>% # can rename in the select step also
  group_by(penguin_species, sex) %>% 
  summarize(mean_body_mass_g = mean(body_mass_g, na.rm = TRUE), 
            mean_flipper_length_mm = mean(flipper_length_mm, na.rm = TRUE),
            count = n())
```

## Some new stuff

A few variations on dplyr::across()

Create a summary table of mean values by species for all numeric variables in the penguins dataset

```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(across(where(is.numeric), mean, na.rm = TRUE))
```

Calculate the mean value of all columns that start with "bill"

```{r}
penguins %>% 
  group_by(species, island) %>% 
  summarize(across(starts_with("bill"), mean, na.rm = TRUE))
```

Write code that will return the minimum value of any columns that end with the character "mm", grouped by island and year

```{r}
penguins %>% 
  group_by(year, island) %>% 
  summarise(across(ends_with("mm"), min, na.rm = TRUE))
```

Find and return both the mean and standard deviation of all columns starting with "bill", grouped by species, and returning a new name that combines the original column with the function.

```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(across(starts_with("bill"), 
                   list(mean = mean, sd = sd), 
                   na.rm = TRUE, 
                   .names = "{.col}_{.fn}"))
```

```{r}
penguins_table <- penguins %>% 
  group_by(species, sex) %>% 
  summarize(mean_mass = mean(body_mass_g, na.rm = TRUE),
            sd_mass = sd(body_mass_g, na.rm = TRUE)) %>% 
  kable(col.names = c("Species", "Sex", "Mean body mass (g)", "SD body mass (g)")) %>% 
  kable_styling(full_width = FALSE)
```

```{r}
penguins_table
```

```{r}
chart1 <- penguins %>% 
  ggplot(aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point()
```

```{r}
chart2 <- penguins %>% 
  ggplot(aes(x = bill_length_mm)) +
  geom_histogram()
```

```{r}
# when patchwork is loaded, can combine charts following PEMDAS
(chart1 + chart2) / (chart2 + chart1)
```

```{r}
#| fig-cap: Penguin body mass observations by species.
#| fig-alt: Jitterplot showing the body mass measurements (y axis, in grams) for three penguin species (Adelie, Gentoo, and Chinstrap on x-axis) observed near Palmer Archipelago, Anartica. Adelie and Chinstraps have similar body mass distributions, with Gentoos noticeably larger than both
ggplot(penguins, aes(x = species, y = body_mass_g)) + geom_jitter(width = 0.2, alpha = 0.5, color = "navy")
```

## Reprex!

```{r}
library(tidyverse)

ggplot(data = mtcars, aes(x = mpg, y = hp)) + geom_point()

# select and copy to clipboard above (what you want to use a reporducible example) then type "reprex()" in the console and run
```
